---
title: "ecp: An R Package for Nonparametric Multiple Change Point Analysis of Multivariate Data"
author: "Joshua Kunst"
date: "Friday, January 23, 2015"
output:
  html_document:
    keep_md: yes
---

Taken from [ecp: An R Package for Nonparametric Multiple Change Point Analysis of Multivariate Data](http://www.jstatsoft.org/v62/i07/paper).

Basically, give a numeric ordered matrix (by column), the function `e.divisive` estimate the index where are change points in the distribution.


```{r}
set.seed(250)
library("ecp")

p1 <- rnorm(100, 0, 1) # N(0,1)
p2 <- rnorm(100, 0, 3) # N(0,3) # change variance
p3 <- rnorm(100, 2, 3) # N(2,3) # change mean
p4 <- rnorm(100, 2, 4) # N(2,4) # change variance

# The parameter must be a matrix

Xnorm <- matrix(c(p1, p2, p3, p4), ncol = 1)

output0 <- e.divisive(Xnorm) # take some time 3 secs in a I3

output0


output0$estimates
output1$k.hat
output1$order.found

```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
