---
output:
  html_document:
    highlight: null
    keep_md: yes
    theme: null
---
```{r, echo=FALSE}
library(knitr)
opts_chunk$set(fig.path="plot/gtfs-", fig.align="center", dpi = 85,
               cache=TRUE, cache.path="cache/",
               message=FALSE)
suppressWarnings(dir.create("plot"))
```

Recently an study says that Santiago, city where I live, has one of the best public transport system in LATAM (WAT?! define *best* please!). So I've search for some information and I found [this](http://www.siemens.com/press/pool/de/feature/2014/infrastructure-cities/2014-06-mobility-opportunity/slide-credo.pdf#page=6). Anyway I tried to find some related data/information to work and I found the *Transantiago GTFS*. GTFS stands from *General Transit Feed Specification* and is a format for public transportation schedules and geographic data.

()

Recently an study  

Usually is a zip file with txt files: `routes.txt`, `shapes.txt`, `stations.txt`, etc. And define  


```{r}
library(ggplot2)
shapes <- read.csv("data/shapes.txt")
```

The content is the 

```{r, results='asis', echo=FALSE}
kable(head(shapes))
```

So, it's ultra simple plot with ggplot.

```{r, plot_simple}
p <- ggplot(shapes) +
  geom_path(aes(shape_pt_lon, shape_pt_lat, group=shape_id), size=.2, alpha=.1) +
  coord_equal()
p
```

You can also embed plots, for example: You can also embed plots, for example: You can also embed plots, for example: You can also embed plots, for example: You can also embed plots, for example: You can also embed plots, for example: You can also embed plots, for example: You can also embed plots, for example: 

```{r}
library(devtools)
library(plyr)
library(dplyr)
source_url("https://raw.githubusercontent.com/jbkunst/reuse/master/R/gg_themes.R")
```

You can also embed plots, for example: You can also embed plots, for example: You can also embed plots, for example: You can also embed plots, for example: You can also embed plots, for example: You can also embed plots, for example: You can also embed plots, for example: You can also embed plots, for example: 


```{r}
routes <- read.csv("data/routes.txt")
trips <- read.csv("data/trips.txt")
stops <- read.csv("data/stops.txt")

stops_metro <- stops %>% filter(!grepl("\\d", stop_id))

shapes_colors <- left_join(left_join(shapes %>% select(shape_id) %>% unique(),
                                     trips %>% select(shape_id, route_id) %>% unique(),
                                     by="shape_id"),
                           routes %>% select(route_id, route_color) %>% unique(),
                           by = "route_id")

shapes_colors <- shapes_colors  %>% mutate(route_color = paste0("#", route_color))

routes_metro <- routes %>% filter(grepl("^L\\d",route_id))

shapes_metro <- shapes %>% filter(shape_id %in% trips$shape_id[trips$route_id %in% routes_metro$route_id])

shapes_colors_metro <- shapes_colors %>% filter(shape_id %in% trips$shape_id[trips$route_id %in% routes_metro$route_id])
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot. Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot. Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot. Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot. 


```{r, plot_complex, dev.args=list(bg="black"), fig.height=7.5}
p2 <- ggplot() +
  geom_path(data=shapes, aes(shape_pt_lon, shape_pt_lat, group=shape_id),
            color="white", size=.2, alpha=.1) +
  geom_path(data=shapes_metro, aes(shape_pt_lon, shape_pt_lat, group=shape_id, colour=shape_id),
            size = 2, alpha=.7) +
  scale_color_manual(values=shapes_colors_metro$route_color) +
  geom_point(data=stops_metro, aes(stop_lon, stop_lat),
             shape=21, colour="white", alpha =.8) +
  coord_equal() +
  theme_null() +
  theme(plot.background = element_rect(fill = "black", colour = "black"),
        title = element_text(hjust=1, colour="white", size = 8),
        axis.title.x = element_text(hjust=0, colour="white", size = 7)) +
  ggtitle("TRANSANTIAGO\nSantiago's public transport system") + 
  xlab(sprintf("Jkunst.com %s",format(Sys.Date(), "%Y")))
p2
```

You can also export the image in a high resolution pds like [this](https://github.com/jbkunst/r-posts/blob/master/GTFS-Transantiago-ggplot2/plot/gtfs-transantiago.pdf?raw=true) with:

```{r}
ggsave(filename = "plot/gtfs-transantiago.pdf", plot = p2, width = 16, height = 12, bg = "black")
```




